---
title: " "
output: html_notebook
---

```{r git, include=FALSE}
use_git_config(user.name="ktemadarko",user.email="k.temadarko@gmail.com")
```

### Factors associated with early birth

#### Using data from DHS Ghana 2014

#### By Korantema Darko

```{r load data, include=FALSE}
#library(haven, tidyverse,tidylog,ggforce,ggpubr,scales)
idhs <- read_dta("idhs_00006.dta/idhs_00006.dta")
idhs<-idhs[-c(1:9)]

```

```{r data cleaning, include=FALSE}
#dat1<-tibble(idhs)%>%
  #select(c(10,19:60))%>%
  #filter( wkearnsmore<7)

#fp<-idhs%>%
  filter(ageat1stbirth<99,resident<2,agefrstmar<99,currwork<98,wkmaternity<99,
         wkearnsmore<9)%>%
  select(-starts_with("rel"),-starts_with("age5"))%>%
  mutate(across(c(urban,resident,marstat,currwork, wkmaternity,wkearnsmore), as_factor))
         
#fp1<-idhs%>%
  filter(ageat1stbirth<99,resident<2,currwork<98)%>%
  select(-starts_with("rel"),-starts_with("age5"))%>%
  mutate(across(c(urban,resident,marstat,currwork), as_factor))

fp2<-idhs%>%
  filter(ageat1stbirth<99,resident<2,agefrstmar<99,lit2<98)%>%
  select(-starts_with("rel"),-starts_with("age5"))%>%
  mutate(across(c(urban,resident,marstat,currwork, wkmaternity,wkearnsmore,wealthq,31:34,36:83), as_factor))%>%
  mutate(across(c(urban,resident,marstat,currwork, wkmaternity,wkearnsmore,wealthq,31:34, 36:83,lit2,educlvl,decbighh), droplevels))
#%>%
  #add_column(cut(fp2$agefrstmar,breaks=c(-Inf,19,Inf), labels=c("married before 19","married after 19")),.after="agefrstmar")


fp2$early_mar=cut(fp2$agefrstmar,breaks=c(-Inf,19,Inf), labels=c("married before 19","married after 19"))

fp2$early_bir=cut(fp2$ageat1stbirth,breaks=c(-Inf,19,Inf), labels=c("birth before 19","birth after 19"))
```

```{r Table1, echo=FALSE}
table1(~age+early_mar+lit2+educlvl+edyrtotal|early_bir,droplevels=T, data=fp2)
```

```{r boxplot, echo=FALSE}
ggplot(fp2,aes(ageat1stbirth))+geom_histogram(colour="white")+scale_x_continuous(n.breaks=35)+ggtitle("A histogram showing the distribution of age of 1st birth")+
  gghighlight(ageat1stbirth<19)

p=ggplot(fp2,aes(fill=urban))+coord_flip()

#p+geom_histogram(color="grey")+scale_x_continuous(n.breaks=12)+ggtitle("A histogram showing the distribution of age of 1st birth")+facet_grid(urban~early_mar)+scale_x_continuous(breaks=pretty_breaks())
#p+geom_density(alpha=0.2, fill="#aabbcc")


#p+geom_boxplot(aes(x=ageat1stbirth))+ggtitle("A boxplot showing the distribution of age of 1st birth by urban/rural")+facet_wrap(~urban)

#p+geom_boxplot(aes(x=ageat1stbirth))+ggtitle("A boxplot showing the distribution of age of 1st birth by urban/rural")+facet_wrap(~early_mar)

#ggplot(fp2,aes(x=ageat1stbirth,fill=urban))+geom_histogram(position = "identity",color="grey")+scale_x_continuous(n.breaks=40)+ggtitle("A boxplot showing the distribution of age of 1st birth by urban/rural")

ggplot(fp2,aes(x=ageat1stbirth,color=urban,fill=urban))+geom_histogram(position = "dodge")+scale_x_continuous(n.breaks=40)+ggtitle("A histogram showing the distribution of age of 1st birth by urban/rural")


#ggplot(fp2,aes(x=ageat1stbirth,color=urban,fill=urban))+geom_histogram(position = "identity",color="grey")+scale_x_continuous(n.breaks=40)+ggtitle("A histogram showing the distribution of age of 1st birth by urban/rural")

#ggplot(fp,aes(x=ageat1stbirth,color=urban,fill=urban))+geom_bar(position = "dodge",color="grey")+scale_x_continuous(n.breaks=40)+ggtitle("A histogram showing the distribution of age of 1st birth by urban/rural")+coord_flip()

#+geom_label(stat="count",aes(label=..count..))
```

```{r age  birth versus urban, echo=FALSE}
facet=ggplot(fp2,aes(x=ageat1stbirth,color=urban,fill=urban))+geom_bar(position = "dodge",color="grey")+scale_x_continuous(breaks=pretty_breaks())+ggtitle("A histogram showing the distribution of age of 1st birth by urban/rural")

mar=ggplot(fp2,aes(x=agefrstmar,color=urban,fill=urban))+geom_bar(position = "dodge",color="grey")+scale_x_continuous(breaks=pretty_breaks())+ggtitle("A histogram showing the distribution of age of 1st marriage by urban/rural")
#mar+gghighlight( ageat1stbirth<19,use_direct_label = FALSE)+facet_zoom(xlim = c(10,20))

facet+facet_zoom(xlim = c(11,14))

facet+facet_zoom(xlim = c(15,25))

facet+facet_zoom(xlim = c(26,34))
```

```{r ggplot setup, include=FALSE}
b=ggplot(fp2)+scale_x_continuous(n.breaks = 20)
m=ggplot(fp2)+scale_x_continuous(n.breaks = 20)
c=ggplot(fp2)+scale_x_continuous(n.breaks = 20)
emb=ggplot(fp2)+scale_x_continuous(n.breaks = 20)
```

```{r age_birth va age_married, include=FALSE}
#showing under 19 years at first birth
#age=b+geom_bar(aes(x=ageat1stbirth ))+scale_y_continuous(n.breaks = 10)
#age+gghighlight( ageat1stbirth<19,use_direct_label =FALSE)+facet_wrap(~urban,scales="free",nrow=2)

#showing under 19 years at first birth
#m+geom_bar(aes(x=agefrstmar, fill=urban ))+gghighlight( ageat1stbirth<19,use_direct_label = FALSE)+facet_wrap(~urban,nrow=2, scales="free")

b+geom_bar(aes(x=ageat1stbirth))+gghighlight( cheb>5,use_direct_label = FALSE)
chb+gghighlight(cheb>5,use_direct_label = FALSE)+facet_wrap(~urban,scales="free")
```

```{r computation dpylr, include=FALSE}
a=fp2%>%select(ageat1stbirth)%>%
  count(ageat1stbirth<19)%>%
  mutate(round((n/5897)*100))#%>%
a
t1=as.character(a[2,3])

curwk=fp2%>%select(currwork)%>%
  count(currwork=="no")%>%
  mutate(round((n/5897)*100))#%>%
curwk

emar=fp2%>%select(agefrstmar)%>%
  count(agefrstmar<19)%>%
  mutate(round((n/5897)*100))#%>%
emar

marbir19<-idhs%>%
  filter(ageat1stbirth<19,agefrstmar<19)%>%
  mutate(across(c(currwork),as_factor))

curwk1=marbir19%>%select(currwork)%>%
  count(currwork=="no")%>%
  mutate(round((n/5897)*100))#%>%
curwk1

```

```{r plots, echo=FALSE}
mar=c+geom_bar(aes(x=agefrstmar))

ab=gghighlight( ageat1stbirth<19,use_direct_label = FALSE)#+
ab1=ggtitle(paste("highlight birth <=18 percent of total",t1,"%"))

crw=gghighlight( currwork=="no",use_direct_label = FALSE)
crw1=ggtitle(paste("highlight not currently working n=",curwk[2,3],"%"))
em=gghighlight( agefrstmar<19,use_direct_label = FALSE)
em1=ggtitle(paste("highlight married before 19 n=",emar[2,3],"%"))

mar+ab+ab1
#mar+crw+crw1
mar+em+em1
#em2=em+em1
#+facet_wrap(~urban,nrow=2, scales="free")

#showing under 19 years at first marriage

#showing more than 6 children
chbg=geom_bar(aes(x=cheb))
chb=c+chbg+scale_y_continuous(n.breaks = 10)
#chb+gghighlight(max(cheb))

#c+geom_bar(aes(x=cheb,fill=urban), position="dodge")+scale_y_continuous(n.breaks = 10)

c_dodge=c+geom_bar(aes(x=cheb), position="dodge")
c_dodge+ab+ab1
c_dodge+em+em1
#c_dodge+crw+crw1

#emb1=emb+chbg+crw+crw1
##emb1

#emb+chbg+em
#cheb+facet_wrap(urban~.)
```

```{r include=FALSE}
cheba=c+geom_col(aes(x=ageat1stbirth,y=cheb))+scale_y_continuous(n.breaks = 20)+coord_flip()
c+geom_col(aes(x=ageat1stbirth,y=cheb))+scale_y_continuous(n.breaks = 20)+gghighlight(cheb>5)

cheba+gghighlight(cheb>5) #does not show relevance geom_point

#c+geom_point(aes(y=ageat1stbirth,x=cheb))+scale_y_continuous(n.breaks = 20)+gghighlight(agefrstmar<19)
c+geom_col(aes(y=ageat1stbirth,x=cheb))+scale_y_continuous(n.breaks = 20)+gghighlight(agefrstmar<19)

cheba+gghighlight(agefrstmar<19)
#cheba+gghighlight(cheb>5,label_key = agefrstmar)
cheba+gghighlight(ageat1stbirth<19)
cheba+gghighlight(currwork=="no")
cheba+facet_wrap(~urban)

#marbir=m+geom_col(aes(x=ageat1stbirth,y=agefrstmar))+scale_y_continuous(n.breaks = 20)
#marbir+gghighlight(agefrstmar<19)
cheba+gghighlight(agefrstmar<19)+facet_wrap(~urban)

agebirth_cheb<-fp1%>%
  tidycomm::crosstab(cheb,ageat1stbirth,urban)
agebirth_cheb
```

```{r tests, include=FALSE}
fp2 %>%
  tidycomm::crosstab(urban,early_mar, add_total = TRUE)

fp2 %>%
  tidycomm::crosstab(wealthq,early_mar, percentages=TRUE)


fp2 %>%
  tidycomm::crosstab(urban,early_bir)

fp2 %>%
  tidycomm::crosstab(urban,early_mar)

fp2 %>%
  tidycomm::crosstab(urban,early_mar,wealthq)

fp2%>%group_by(wealthq)%>%count(early_bir,early_mar,urban)
#rename(fp2$early_mar=fp2$cut(...))
fp2%>%group_by(wealthq)%>%count(early_bir,early_mar)

fp2%>%group_by(wealthq)%>%count(early_bir,early_mar,urban)

fp2%>%group_by(wealthq)%>%count(early_bir,early_mar,currwork)

fp2%>%group_by(early_bir)%>%count(early_mar)



```

Ho=The relationship between early birth and early marriage is independent HA=The relationship between early birth and early marriage is dependent since p\<0.05 reject the Ho, they are dependent

```{r association, include=FALSE}
t1<-fp2%>%
  tidycomm::crosstab(early_bir,early_mar, percentages=TRUE,add_total = TRUE)



t4<-fp2 %>%
  tidycomm::crosstab(early_bir,early_mar,add_total = TRUE)
t5<-fp2 %>%
  tidycomm::crosstab(early_bir,wealthq,add_total = TRUE)
ebwq1<-fp2%>%group_by(wealthq)%>%count(early_bir)

t2<-fp2%>%
  tidycomm::crosstab(early_bir,wealthq, percentages=TRUE,add_total = TRUE)
t6<-fp2%>%
  tidycomm::crosstab(wealthq,early_bir, percentages=TRUE,add_total = TRUE)
t6
t2

#ebwq
#ebwq1
t2


t3=tabyl(fp2,early_mar,early_bir)
t4=tabyl(fp2,early_mar,early_bir)

tr1<-xtabs(~early_bir+early_mar,fp2)
tr2<-xtabs(~wealthq+early_bir,fp2)
tr3<-xtabs(~early_bir+wealthq,fp2
           )
chisq<-janitor::chisq.test(tr1)

t1
#tidy(t3)
#stats::chisq.test(t3)
chisq
tidy(chisq)
chisq$observed
chisq$expected
chisq$residuals

#chisq1<-chisq.test(ebwq1, tabyl_results= TRUE)
#chisq1$observed
#chisq1$expected
#chisq1$residuals

t1
t2
t3
t4

tr1<-xtabs(~early_bir+early_mar,fp2)
tr2<-xtabs(~wealthq+early_bir,fp2)
tr3<-xtabs(~early_bir+wealthq,fp2
           )
t5=tibble(fp2$early_mar,fp2$early_bir)
#t5
tab2by2.test(x=fp2$early_mar, y=fp2$early_bir)
tab2by2.test(tr1)

loddsratio(tr1)

#mosaic(t1)
assoc(t1)
#assocplot(t1)
epitools::epitable(x=fp2$early_mar, y=fp2$early_bir)
epitools::epitab(x=fp2$early_mar, y=fp2$early_bir)

#lit2
#fp2$educlvl
#edyrtotal
#decbighh


tr1
tr2
tr3
mosaic(tr1)
t4
t1
#assoc(tr1)
#assocplot(tr1)
chisq
chisq$observed
 
mosaic(tr2)
#mosaic(tr3)
t5
t2
#t6
assoc(tr2)

assocplot(tr2)
```

```{r one}
#tr2
#tr3
mosaic(tr1)
t4
t1

chisq
chisq$observed
 
mosaic(tr2)
#mosaic(tr3)
t5
t2
```

\`\`\`
